\chapter{Introduction}

\section{The Standard Model}
\label{SM}
For the last 50 years particle physicists have very successfully described the short distance interactions of elementary particles using the Standard Model.  
The Standard Model consists of two quantum field theories: the Electroweak Theory and \gls{QCD}, which describe the interactions of 12 fundamental spin $\frac{1}{2}$ fermions.  
While the Standard Model contains no explanation for gravity, dark matter, dark energy, etc. it does remain the most successful model available to describe reality on the smallest scales and at the highest energies.  

\begin{table}
  \centering
  \begin{tabular}{ |c|c|c|c|c|c|}
  \hline
   fermion & mass [GeV] & spin & electric charge & colour charge & generation \\
  \hline
  \multicolumn{6}{|c|}{charged leptons} \\
  \hline
    e    & 5.11 x$10^{-4}$ & $\frac{1}{2}$ & -1 & no & 1 \\
  $\mu$  & 1.06 x$10^{-1}$  & $\frac{1}{2}$ & -1 & no & 2 \\
  $\tau$ & 1.78            & $\frac{1}{2}$ & -1 & no & 3 \\
  \hline
  \multicolumn{6}{|c|}{neutral leptons} \\
  \hline
  $\nu_e$      & 0 & $\frac{1}{2}$ & 0 & no & 1 \\
  $\nu_{\mu}$  & 0 & $\frac{1}{2}$ & 0 & no & 2 \\
  $\nu_{\tau}$ & 0 & $\frac{1}{2}$ & 0 & no & 2 \\
  \hline
  \multicolumn{6}{|c|}{up-type quarks} \\
  \hline
  u & 2.3 x$10^{-3}$ & $\frac{1}{2}$ & $+\frac{2}{3}$ & yes & 1 \\ 
  c & 1.28           & $\frac{1}{2}$ & $+\frac{2}{3}$ & yes & 2 \\
  t & 173.5          & $\frac{1}{2}$ & $+\frac{2}{3}$ & yes & 3 \\
  \hline
  \multicolumn{6}{|c|}{down-type quarks} \\
  \hline
  d & 4.8 x$10^{-3}$ & $\frac{1}{2}$ & $-\frac{1}{3}$ & yes & 1 \\
  s & 9.5 x$10^{-2}$ & $\frac{1}{2}$ & $-\frac{1}{3}$ & yes & 2 \\
  b & 4.18           & $\frac{1}{2}$ & $-\frac{1}{3}$ & yes & 3 \\
  \hline
  \end{tabular}
  \caption[Properties of known spin-$\frac{1}{2}$ bosons in the Standard Model.]
        {\small Properties of the known spin-$\frac{1}{2}$ fermions in the Standard Model~\cite{PDG}.  The quark masses have been estimated using the $\overline{\mathrm{MS}}$ renormalization scheme at a scale $\mu = 2$ GeV, and while the neutrino masses are non-zero, they are small enough that they are approximated as zero for the purposes of this thesis.}
\label{table:Fermions}
\end{table}


The fermions, listed in Table~\ref{table:Fermions}, are categorized into three generations of four particles, with each generation being a heavier copy of the previous one.  
Each generation includes two quarks, particles that have colour charge and are therefore subject to the strong force described by \gls{QCD}, and two leptons that have no colour charge.  
The six types (flavours) of quarks are organized into generational pairs, the up (u) and the down (d), the charm (c) and the strange (s), and the top (t) and the bottom (b).  
Similarly each generation of leptons is composed of one charged lepton, known as the electron (e), the muon ($\mathrm{\mu}$), and the tau ($\mathrm{\tau}$), and their neutral counterparts the neutrinos ($\nu_{\mathrm{e}}, \nu_{\mu}, \nu_{\tau}$).  
In addition to each of these matter particles there is a complimentary anti-particle, which has the opposite quantum numbers, e.g. charge, to it's regular counterpart and is denoted using an over bar.  
It is worth noting that the entirety of the periodic table of elements, and therefore all regular matter, require only the first generation of particles to construct.  
The Standard Model also includes spin 1 bosons which mediate the various forces included in the theory (see Table~\ref{table:Bosons}). 

\begin{table}
  \centering
  \begin{tabular}{ |c|c|c|c|c|c|}
  \hline
  interaction     & boson             & mass [GeV] & spin & electric charge &  colour charge \\ \hline
  \multicolumn{6}{|c|}{force carrying bosons} \\ \hline
  electromagnetic & $\gamma$ (photon) & 0          & 1 & 0 & no \\ \hline
  \multirow{2}{*}{Weak} & $W^{\pm}$ & 80.39 & 1 & $\pm1$ & no  \\ 
                        & Z & 91.19 & 1 & 0 & no \\ \hline
  strong & g (gluon) & 0 & 1 & 0 & yes \\ \hline 
  \multicolumn{6}{|c|}{non-force carrying bosons} \\ \hline
  --- & Higgs & 125.09 & 0 & 0 & no \\ \hline
  \end{tabular}
  \caption[Properties of known bosons in the Standard Model.]
        {\small Properties of known bosons in the Standard Model.}
\label{table:Bosons}
\end{table}



The electroweak interaction has four force carrying bosons: the photon, the Z, and W$^{\pm}$.  
While the electroweak force is unified at high energies, at lower every day energies the electroweak symmetry is spontaneously broken into the electromagnetic force and the weak force.  
In the Standard Model this breaking is explained by introducing a new scalar field with a non-zero vacuum expectation value, known as the Higgs field.  
This breaking into two separate forces also allows the W and Z bosons to acquire their very large observed mass through the Brout-Englert-Higgs (BEH) mechanism.  
The large mass of the W and Z bosons also leads to the relatively short range of the weak interaction.  
This Higgs field also allows the fermions in the Standard Model to obtain mass as well.  
A measurable consequence of this theory is the presence of a massive spin-0 boson called the Higgs Boson, which was discovered in 2012 at the \gls{LHC}.  

As previously mentioned, \gls{QCD} describes the strong force, which affects particles carrying colour charge much like the electric force affects particles carrying an electric charge.  
Unlike electromagnetism which only has one charge, the electric charge, \gls{QCD} has three colour charges known as red, green, and blue.  
Another notable difference between the two interactions is that while the force mediator of the electric force (the photon) has no electric charge itself, this is not the case for the force carriers of the strong interaction (gluons).  
One important consequence of this difference is that the strength of the strong force does not weaken with increasing distance between the particles in question.  
This difference means that quarks cannot be found as individual particles but come in bound ``colour neutral'' states, a phenomenon known as confinement.  
While these bound states are colour neutral a much weaker residual force still exists, and while it decreases quite rapidly with distance it is this residual force which is responsible for holding protons and neutrons together in the nuclei of atoms.  

These bound states contain an infinite number of quark/anti-quark pairs constantly being created and annihilated (called sea quarks) and gluons, but the overall identity and properties of a given bound state are determined by it's valence quarks.  
Bound states can have either three valence quarks (or anti-quarks) which are known as the baryons, or a quark/anti-quark pair, called the mesons.  
Baryons and mesons are collectively known as hadrons.  
As hinted at above, familiar examples of these bound states include protons (uud) and neutrons (udd).  
As a result of these complicated bound states the masses of the quarks cannot be measured directly, but must be approximated by measuring the bound states and making certain theoretical assumptions~\footnote{The t quark is the exception, as it decays fast enough that no bound states are formed.}.
 

\section{Experimental Particle Physics}
\label{Sec:Experi}
The Standard Model and all of its predictions can be studied in a variety of settings.  
One category of experiment involves accelerating stable, or long-lived, particles and bringing them to collision inside of a detector, allowing for the study of the various particles of the Standard Model that are created in the collisions.  
The particles being accelerated can be both elementary (electrons for example) or composite (like protons), with the different choices affecting the types of collisions that will be observed.  
A relevant example of such a detector/accelerator combination is \gls{ATLAS} at the \gls{LHC}.  

The \gls{LHC} is a machine that straddles the border of Switzerland and France at the \gls{CERN} in Geneva.  
While it can produce heavy ion collisions, it is best known for its ability to produce higher energy proton-proton collisions than any other accelerator facility in the world.   
Bringing protons from rest to within a few meters per second from the speed of light is a multi-step process, a description of which can be found in the LHC Design Report~\cite{LHCDesignReport} while a summary can be found in section~\ref{Sec:LHC}.  

\gls{ATLAS} is a general purpose detector that has been placed in the LHC accelerator ring to observe collisions along with three other detectors.  
There is another general purpose detector known as the \gls{CMS}, in addition to two more specialized detectors called \gls{LHCb} and \gls{ALICE}.  
Detectors like \gls{ATLAS} and \gls{CMS} are designed to be able to accurately measure a wide variety of processes allowing them to more fully explore the collisions provided by the \gls{LHC} to push the Standard Model to its limit and search for potential new physics beyond the Standard Model.  
Details on how this is accomplished in ATLAS will be provided in section~\ref{Sec:ATLAS}.  

As the \gls{LHC} collides protons, and therefore quarks and gluons, the most commonly produced objects in the \gls{ATLAS} detector are collimated sprays of particles called jets.  
Jets are the result of an interplay between a high energy colour charged particle being ejected from a collision and the confinement property of the strong force allowing only colour neutral particles to exist.  
The formation of jets will be discussed in greater detail in Sec.~\ref{jets}.  
With this large abundance of jets they are inevitably a signature for a large number of final states being searched for or studied in ATLAS, with many requiring a very accurate energy and momentum calibration for the jets.  
This thesis presents detailed studies of a technique for measuring the jet energy scale \textit{in-situ} using events where jets are produced along with other well measured objects (photons, electrons and muons).  

\section{Units and Conventions}
 
For the remainder of this thesis it will be worth keeping a few particle physics conventions in mind.  
The most commonly used is the non-SI unit of energy, the \gls{eV}, which is the amount of energy a singly-charged particle gains after being accelerated across a potential of 1 Volt.  
One eV is the equivalent of 1.602$\times 10^{-19}$ Joules.  
Energies relevant to this thesis are typically be much larger than a single eV, so gigaelectronvolt (GeV, $10^9$ eV) and teraelectronvolts  (TeV, $10^{12}$ eV) are more commonly used.  
Traditionally electronvolts are also used to describe mass and momentum.  
This is made possible using another particle physics convention which is to set the speed of light c to be unitless and equal to one, allowing mass to be described in terms of eV/c$^2$ (see tables~\ref{table:Fermions} and \ref{table:Bosons}), and momentum to be referenced in terms of eV/c.  
It is also standard practice to treat $\hbar$ the same way.

