%
%  LaTeX source:
%  This is a skeleton thesis, provided by (?) Wolfgang Meuck
%  to be copied, along with thesisstyle.sty to your computer.
%
%  Slightly modified by Andrew Kurn, May 2009
%
\documentclass[12pt,letterpaper,oneside]{book}
% Use either 11pt or 12pt! 10pt is too small, although supported. 

\usepackage{epsfig}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\usepackage{array}
\usepackage{booktabs,amsmath}
\usepackage{makecell}
\usepackage{rotating}
\usepackage{multicol}
\usepackage{slashed}
%\usepackage{slashbox}
\usepackage{float}
\usepackage{amsmath}
%\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{needspace}
\usepackage{afterpage}
\usepackage[sort&compress]{natbib}
\usepackage[titletoc]{appendix}

% Load the package
\usepackage[acronym]{glossaries} 
% Generate the glossary
\makeglossaries
\usepackage[xindy]{imakeidx}
\makeindex


\include{Abbreviations}
 

 


\usepackage[headings]{thesisstyle}
% Options plain and headings are supported. 

% This makes use LaTeX's new font selection scheme (nfss).
\usepackage[T1]{fontenc}

% You might consider using postscript fonts.
% `Times' can be included used this: (2nd line is for math fonts)
\usepackage{times}
%\usepackage{mathptm}  
\newcommand{\MET}{\ensuremath E_T^{miss}}
\newcommand{\vMET}{\ensuremath \vec{E}_T^{miss}}
\newcommand{\PT}{\ensuremath p_{\mathrm {T}}}
\newcommand{\vPT}{\ensuremath \vec{p}_{\mathrm {T}}}
% The following packages might be useful for formula typesetting and
% for the inclusion of graphs, respectively.
%\usepackage{amsmath,amssymb}

%Turned off by Jamie on the 13th of July, 2016
%\usepackage{graphicx}

% Substitute your information 
\author{Arthur James Horton}
\qualifications{BSc., University of Prince Edward Island 2010}
\qualifications{MSc., Simon Fraser University 2013}
\title{IN-SITU MEASUREMENT OF THE JET ENERGY SCALE AND STUDIES OF JET STRUCTURE AT ATLAS}
\degree{Doctor of Philosophy}
\department{Physics \\ Faculty of Science}
\submitdate{????}
%\thesistype{thesis}   % standard is "thesis"
\entity{Department}   % standard is "Department"
%\copyrightyear{}      % standard is current

% If you want to grant the copyright instead of reserving it,
% uncomment the next line!
%\grantcopyright   

% Insert here the members of your examining committee 
\chair{Dr. Eldon Emberly (Chair)\\
Associate Professor}
\committee{Dr. Michel Vetterli, Senior Supervisor\\
Professor}
\committee{Dr. Dugan O'Neil, Supervisor\\
Professor}
\committee{Dr. Bernd Stelzer, Supervisor\\
Associate Professor}
%\committee{Dr. Corina Andreoiu, Internal Examiner\\
%Assistant Professor\\
%Department of Chemistry}

\approvaldate{22 March, 2013}

\abstract{
This thesis presents results for the determination of the ATLAS jet energy scale (JES) using the Missing $E_{\mathrm T}$ Projecting Fraction (MPF) method along with studies to better understand and validate the MPF.
Hadronic jets are the most commonly observed objects in proton-proton collisions, and are therefore a part of most final states for processes which are studied at the Large Hadron Collider (LHC).
The abundance of jets makes a precise knowledge of the JES essential to the success of the ATLAS physics program.
This thesis uses the MPF in events where either a photon or a Z boson is produced back-to-back with a jet to provide an uncertainty on the response of the calorimeter which is below 1\% for jets between 30 GeV and 1 TeV.

Studies measuring the impact of the underlying event on the MPF's ability to measure the response of the hadronic recoil are also presented, which validate the previously held assumption that the MPF is insensitive to these effects.  
In addition, studies into the relation between the measured recoil response and the desired jet response are presented.  
This include measures of the flow of energy across the jet boundary during the showering process and the effect of low energy/low response particles near the fringe of the recoil on the total measured response.  
These measurements show up to a 10\% difference between the jet response and the recoil response for jets reconstructed with the anti-$k_\mathrm{t}$ algorithm with midrange size parameters (0.4-0.7).  
These differences however show little dependence on physics modeling choices (less than 1\%), on which the Monte Carlo jet calibration is based. 
These results put the MPF technique on a firmer ground, and they will reduce future JES uncertainties for jets with energies below 100 GeV.  
}


\begin{document}
\frontmatter          % switches to roman page numbering
\maketitle

%\chapter[Dedication]{}  % writes `Dedication' into table of contents,
			% but not on actual page 
\begin{center}
%\Large
%\emph{\`A tous ceux que j'aime:}
%\\
%\emph{Mum, Dad, Amy, M\'{e}lanie, Justin, Eli, M\'{e}m\`{e}re, P\'{e}p\`{e}re, Ron, et sp\'{e}cialement Ryan.}
\end{center}

% end of dedication

\chapter{Acknowledgments}
%Words go here


% end of acknowledgements

\newpage
\addcontentsline{toc}{chapter}{\contentsname}
\tableofcontents

\newpage 
\addcontentsline{toc}{chapter}{\listtablename}
\listoftables

\newpage
\addcontentsline{toc}{chapter}{\listfigurename}
\listoffigures

\newpage
\addcontentsline{toc}{chapter}{Glossary}
%Print the glossary
%\glsaddall
\printglossary[type=\acronymtype,nonumberlist]
%\printglossary[title=Acronyms,nonumberlist]
%[type=\acronymtype,title=Abbreviations]

\mainmatter 
% Here comes the content of your thesis. 
% You should not write the text here, but include each chapter
% separately with the \include command.

%\include{chap0}
\include{chap1}
\include{chap2}
\include{chap3}
\include{chap4}
\include{chap5}
\include{chap6}
\include{chap7}
\appendix
\begin{appendices}
%\input{appendixJetRecoBias}
%\input{appendixGJetEM_2016}
%\input{appendixGJetLC_2016}
%\input{appendixZJetEM_2016}
%\input{appendixZJetLC_2016}
\input{appendix_results2016}
\input{appendix_results2015}
\input{appendixJ2_EM}
\input{appendixdPhi_EM}
\input{appendix_Showering}
\input{appendix_ShoweringPhoton}
\input{appendix_Topology}
\input{appendix_TopologyPhoton}
\end{appendices}

\backmatter
\newpage
\include{appendices}


\backmatter
\newpage
\addcontentsline{toc}{chapter}{\bibname}

% Here comes the bibliography

\bibliographystyle{unsrt}
\bibliography{myrefs}

\end{document}
